Color-dual NLSM multi-loop integrands

### 2023-07-24 notes/2loopNLSM.tex

Before proceeding, we note a strange property of the $n$-gon numerator for the pions. In this form, Jacobi relations produce \textit{non-vanishing} values for bubbles on external legs (BELs). However, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\be
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234,l)}-N^{\text{NLSM}}_{(1243,l)}-N^{\text{NLSM}}_{(1342,l)}+N^{\text{NLSM}}_{(1432,l)}
\ee
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $l_i$, we obtain the following expression for the BEL
\be
N^{\text{BEL}}_{1|2,34} = s_{12} (l+k_1)^2 l^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} 
\ee
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\be
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D l}{i\pi^{D/2}} \frac{l^\mu }{l^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\ee
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in large enough dimension, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram. 
\section{Two-loop}
Let's start with a two-loop four-point example. At this order, the basis diagrams for color-dual representations are any two of the Jacobi triple, the double box, the penta-triangle or the cross-box. First we introduce a bit of shorthand notion:
\be
[LR] \equiv  l_L^2 -l_R^2 \qquad [M] \equiv  l_M^2
\ee
Now let's start with the double-box. Below we treat the color legs as $Y$ particles that will get dressed with factors of $[LR]$. Thus, we wish to manifest the following cut:
\be
N^{\text{2box}}_M\equiv \dBox{}{hred}{hred}{hred}{hred}{hred}{hred} = [12][23][34][45][56][61]
\ee
In addition, we want the following cuts to be simultaneously satisfied:
\be
N^{\text{2box}}_R\equiv \dBoxR{hred}{}{hred}{}{hred}{hred}{hred} = [61][12][27][76][4]^2
\ee
\be
N^{\text{2box}}_L\equiv \dBoxL{hred}{hred}{}{hred}{}{hred}{hred}=[45][57][73][34][1]^2
\ee
Where the middle cut takes $l_7^2\rightarrow 0$, the right cut takes $l_3^2,l_5^2\rightarrow 0$ and the left cut takes $l_2^2,l_6^2\rightarrow 0$. Applying these cuts to all the diagrams above, we obtain the following expressions:
\be
\text{Cut}_R(N^{\text{2box}}_M) = \Delta_R + N^{\text{2box}}_R \qquad \text{Cut}_L(N^{\text{2box}}_M) = \Delta_L + N^{\text{2box}}_L
\ee
However, we find that the remainder terms $\Delta_{L/R}$ are proportional to $l_7^2$, and thus vanish on the middle cut:
\be
\text{Cut}_M(\Delta_{L/R})  = 0
\ee
Similarly, applying the complement left/right cuts to the right/left remainder terms, we obtain a single kinematic function:
\be
\text{Cut}_L(\Delta_R) = \text{Cut}_R(\Delta_L) = \Delta_{LR}
\ee
Thus, the double box numerator that simultaneously manifests all of the cut contributing to NLSM are as follows:
\be
\boxed{N^{\text{NLSM}}_{\text{2box}} = N^{\text{2box}}_M - (\Delta_{L}+\Delta_{R})+\Delta_{LR}}
\ee
Applying the cut relations above, we find that this double box numerator factorizes to the desired expressions:
\be
\text{Cut}_{X}(N^{\text{NLSM}}_{\text{2box}} )= N^{\text{2box}}_{X}
\ee
The kinematic expressions obtained for the remainder functions are provided below:
\begin{align}
\Delta_L &= \frac{1}{2}[61][12]([27]+[67])[4]^2[7] 
\\
 \Delta_R &= \frac{1}{2}[34][45]([37]+[57])[1]^2[7]
 \\
\Delta_{LR}&= [1]^2[4]^2[7]^2
\end{align}
While these numerators are not functionally automorphic like the one-loop $n$-gon numerators, they do obey color-kinematics when evaluated on-shell and thus should be valid representations for double copy construction. 
\section{Discussion}\label{sec:Discussion}

>>>>

Before proceeding, we note a strange property of the $n$-gon numerator for the pions. In this form, Jacobi relations produce \textit{non-vanishing} values for bubbles on external legs (BELs). However, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\be
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234,l)}-N^{\text{NLSM}}_{(1243,l)}-N^{\text{NLSM}}_{(1342,l)}+N^{\text{NLSM}}_{(1432,l)}
\ee
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $l_i$, we obtain the following expression for the BEL
\be
N^{\text{BEL}}_{1|2,34} = s_{12} (l+k_1)^2 l^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} 
\ee
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\be
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D l}{i\pi^{D/2}} \frac{l^\mu }{l^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\ee
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in large enough dimension, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram. 
\section{Two-loop}
Let's start with a two-loop four-point example. At this order, the basis diagrams for color-dual representations are any two of the Jacobi triple, the double box, the penta-triangle or the cross-box. First we introduce a bit of shorthand notion:
\be
[LR] \equiv  l_L^2 -l_R^2 \qquad [M] \equiv  l_M^2
\ee
Now let's start with the double-box. Below we treat the color legs as $Y$ particles that will get dressed with factors of $[LR]$. Thus, we wish to manifest the following cut:
\be
N^{\text{2box}}_M\equiv \dBox{}{hred}{hred}{hred}{hred}{hred}{hred} = [12][23][34][45][56][61]
\ee
In addition, we want the following cuts to be simultaneously satisfied:
\be
N^{\text{2box}}_R\equiv \dBoxR{hred}{}{hred}{}{hred}{hred}{hred} = [61][12][27][76][4]^2
\ee
\be
N^{\text{2box}}_L\equiv \dBoxL{hred}{hred}{}{hred}{}{hred}{hred}=[45][57][73][34][1]^2
\ee
Where the middle cut takes $l_7^2\rightarrow 0$, the right cut takes $l_3^2,l_5^2\rightarrow 0$ and the left cut takes $l_2^2,l_6^2\rightarrow 0$. Applying these cuts to all the diagrams above, we obtain the following expressions:
\be
\text{Cut}_R(N^{\text{2box}}_M) = \Delta_R + N^{\text{2box}}_R \qquad \text{Cut}_L(N^{\text{2box}}_M) = \Delta_L + N^{\text{2box}}_L
\ee
However, we find that the remainder terms $\Delta_{L/R}$ are proportional to $l_7^2$, and thus vanish on the middle cut:
\be
\text{Cut}_M(\Delta_{L/R})  = 0
\ee
Similarly, applying the complement left/right cuts to the right/left remainder terms, we obtain a single kinematic function:
\be
\text{Cut}_L(\Delta_R) = \text{Cut}_R(\Delta_L) = \Delta_{LR}
\ee
Thus, the double box numerator that simultaneously manifests all of the cut contributing to NLSM are as follows:
\be
\boxed{N^{\text{NLSM}}_{\text{2box}} = N^{\text{2box}}_M - (\Delta_{L}+\Delta_{R})+\Delta_{LR}}
\ee
Applying the cut relations above, we find that this double box numerator factorizes to the desired expressions:
\be
\text{Cut}_{X}(N^{\text{NLSM}}_{\text{2box}} )= N^{\text{2box}}_{X}
\ee
The kinematic expressions obtained for the remainder functions are provided below:
\begin{align}
\Delta_L &= \frac{1}{2}[61][12]([27]+[67])[4]^2[7] 
\\
 \Delta_R &= \frac{1}{2}[34][45]([37]+[57])[1]^2[7]
 \\
\Delta_{LR}&= [1]^2[4]^2[7]^2
\end{align}
While these numerators are not functionally automorphic like the one-loop $n$-gon numerators, they do obey color-kinematics when evaluated on-shell and thus should be valid representations for double copy construction.

\section{Notes}\label{sec:Notes}

\begin{itemize}
\item ZM and 2D NLSM (PCM) are the same classically but not quantum mechanically.
This observation is due to other people but an easy way to see it.
ZM and 2D NLSM have the same EOM after a field redefinition.
This means that the classical physics is the same but of course the Lagrangians aren't the same and the field redefinition can change the path integral measure.
\item Forcing pions to match ZM in 2D does funny things.
The pion coupling constant is dropped but is implicitly set by the ordered 4pt amplitude which is normalized to be $-k_1 \cdot k_3$ in the package.
I've tabulated what happens when you take a general D NLSM answer and force it to match ZM in 2D.
Let's say that $n_\text{NLSM} = z n_\text{ZM}$ in 2D where I'm matching basis/master numerators.
For 4pt tree, $z=2/3$ (I matched the half ladder).
For 4pt 1-loop, $z$ is free/pure generalized gauge (I matched the box).
For 4pt 2-loop (when you impose certain $\ell$ power counting) $z$ is $216/565=\tfrac{2^3 3^3}{5 \cdot 113}$ (I matched the double box and penta-triangle).
For 6pt tree, $z$ is 12/35 (I matched the half ladder).
So you can match kinematic structure (the combination of Mandelstams) of NLSM to ZM in 2D but you can't get the coupling constant right at loop level, that is, ZM is a different theory quantum mechanically.
Actually these numbers make it look like you can't even match the theories classically but I guess for the theories to match classically you only need the amplitudes to agree and I matched numerators.
\item At 4pt 1-loop $\prod p_L^2 -p_R^2$ agrees with the ZM answer $\prod p_L \eps p_R$ up to some number.
At 6pt 1-loop this is not true and the two answers differ by alternating signs depending on the kinematic configuration.
\item You cannot lift ZM to general D NLSM in the democratic way that Nic described because it fails at 6pt tree.
\item To clean up the NLSM 4pt 2-loop answer I forced the answer to match ZM in 2D.
The ZM answer also only has certain powers of loop momenta $\ell^5$...$\ell^8$ where $\ell$ represents any combination of $\ell_1$ and $\ell_2$.
I enforce this $\ell^5$...$\ell^8$ behavior on the NLSM ansatz too.
\end{itemize}

\section{Discussion}\label{sec:Discussion}
